---
topic: "water"
title: "Trends in Stream and River Flow"
output: 
  pdf_document: envreportutils.internal::print_ver
  html_document:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{=html}
<!--
Copyright 2022 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->
```
```{r packages and options, echo=F, warning=F, results=F, message=F}
library(envreportutils)
library(tidyhydat) # to access hydrometric flow database (GWELLS?)
library(sf) # To work with spatial data (shapefiles, geopackages etc.)
library(leaflet) # To make (interactive) maps
library(leaflegend) # To add legend for station size based on mean flow.
library(EnvStats) # Necessary for Mann-Kendall test
library(data.table) # Necessary for use of fcase().
library(gridExtra) # For making table Grob.
library(ggimage) # For making ggplot obj out of PNG file.
library(tidyverse)
library(ggthemes)
library(ggtext)
library(ggrepel)
library(ggpubr)
library(knitr)
library(knitrProgressBar)
library(shiny) # For leaflet size legend. Hopefully can drop this dependency.

knitr::opts_chunk$set(echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE)

if(!dir.exists('factsheet_figures')) dir.create('factsheet_figures')

# Set options (e.g. page size, how far to look for nearby stations, etc.)

print_page_width = 8.27

print_page_height = 11.69

station_search_dist = 100000

# How many separate bins should be used to display mean flow in leaflet map?
number_size_classes = 4

# What baselayer should we use in the leaflet map?
leaflet_provider_choice = providers$Stamen.Terrain
```

```{r MK_test_as_function}
mk_test = function(dat, variable){
  # browser()
  dat %>% 
    reframe(MK_results = kendallTrendTest(!!sym(variable) ~ Year)[c('statistic','p.value','estimate')]) %>%
    unnest(MK_results) %>%
    unnest_longer(col = MK_results) %>%
    mutate(MK_results_id = c('Statistic','P_value','Tau','Slope','Intercept')) %>%
    pivot_wider(names_from = MK_results_id, values_from = MK_results) %>%
    mutate(trend_sig = fcase(
      abs(Tau) <= 0.05 , "No Trend",
      Tau < -0.05 & P_value < 0.05 & variable %in% c('DoY_50pct_TotalQ','Min_7_Day_DoY'), "Significant Trend Earlier",
      Tau < -0.05 & P_value >= 0.05 & variable %in% c('DoY_50pct_TotalQ','Min_7_Day_DoY'), "Non-Significant Trend Earlier",
      Tau > 0.05 & P_value >= 0.05 & variable %in% c('DoY_50pct_TotalQ','Min_7_Day_DoY'), "Non-Significant Trend Later",
      Tau > 0.05 & P_value < 0.05 & variable %in% c('DoY_50pct_TotalQ','Min_7_Day_DoY'), "Significant Trend Later",
      Tau < -0.05 & P_value < 0.05 & (!variable %in% c('DoY_50pct_TotalQ','Min_7_Day_DoY')), "Significant Trend Down",
      Tau < -0.05 & P_value >= 0.05 & (!variable %in% c('DoY_50pct_TotalQ','Min_7_Day_DoY')), "Non-Significant Trend Down",
      Tau > 0.05 & P_value >= 0.05 & (!variable %in% c('DoY_50pct_TotalQ','Min_7_Day_DoY')), "Non-Significant Trend Up",
      Tau > 0.05 & P_value < 0.05 & (!variable %in% c('DoY_50pct_TotalQ','Min_7_Day_DoY')), "Significant Trend Up"
    )) %>% 
    ungroup()
}
```

```{r load_data}
# Load in data.
annual_flow_dat = readRDS(here::here('app','www/annual_flow_dat.rds'))|>  
  as_tibble()

monthly_flow_dat = readRDS(here::here('app','www/monthly_flow_dat.rds'))|>  
  as_tibble()

hydrograph_dat = readRDS(here::here('app','www/hydrograph_dat.rds'))|>  
  as_tibble()

stations = read_sf(here::here('app','www/stations.gpkg'))

districts = bcmaps::nr_districts() %>% st_transform(crs = 4326)

bc = bcmaps::bc_bound() %>% st_transform(crs = 4326)

```

```{r calculate_constants_and_labels}
# Pretty labels that replace the variable names in the data.
variable_label_df = data.frame(
  varname = c('Average','Min_7_Day','Min_7_Day_DoY','DoY_50pct_TotalQ'),
  labelname = c('Average Flow','7-Day Low Flow','7-Day Low Flow Day-of-year','Date of Freshet')
)

# Most recent data.
most_recent_data = paste(
    monthly_flow_dat %>% filter(Year == max(monthly_flow_dat$Year)) %>% select(Month) %>% distinct() %>% tail(1),
  max(monthly_flow_dat$Year)
)

# Oldest data.
oldest_data = paste(
  monthly_flow_dat %>% filter(Year == min(monthly_flow_dat$Year)) %>% select(Month) %>% distinct() %>% slice(2),
  min(monthly_flow_dat$Year)
)
```

```{r add_most_recent_mean_flow_to_stations}
stations = stations %>%
  left_join(annual_flow_dat |> 
              group_by(STATION_NUMBER) %>%
              slice_max(Year) %>%
              ungroup() %>%
              dplyr::select(STATION_NUMBER,Average,Year_of_Average_Flow = Year))
```

```{r make_leaflet_function}
# Create leaflet map ===========================================================

make_leaflet = function(station, stations){
  
  # Add a column identifying the station of focus of this loop.
  stations_marked = stations %>% 
    mutate(focus_station = STATION_NUMBER == station$STATION_NUMBER) %>% 
    mutate(focus_station = ifelse(focus_station == T, "Station of Interest", "Other Station")) %>% 
    mutate(focus_station = factor(focus_station, levels = c("Station of Interest",
                                                            "Other Station"))) %>%
    # Find other stations within search radius. Use these for labelling.
    st_join(., st_buffer(station,station_search_dist) %>% summarise(within_search_radius = TRUE)) %>% 
    mutate(leaflet_label = ifelse(within_search_radius == T, paste0(STATION_NAME, ", (",STATION_NUMBER,")")))
  
  # For nearby stations that will be labelled, only include the largest station in
  # cases where 2+ stations are quite close (according to the zoom factor of 9 that
  # is currently being used, this might be stations that are within 20 kilometers)
  stations_with_labels = stations_marked %>% 
    filter(!is.na(leaflet_label))
    
  # Add a nested distance matrix for each station.
  stations_with_labels = stations_with_labels %>% 
    left_join(
      st_distance(stations_with_labels, stations_with_labels) %>% 
        as.data.frame() %>% 
        setNames(stations_with_labels$STATION_NUMBER) %>% 
        pivot_longer(cols = everything(), names_to = "STATION_NUMBER", values_to = 'dist_mat') %>% 
        mutate(dist_mat = as.numeric(dist_mat)) %>% 
        group_by(STATION_NUMBER) %>% 
        nest() %>% 
        ungroup(),
      by = join_by(STATION_NUMBER)
    )
    
  # Find which stations are close to which other stations... if within 20 km,
    # just keep label for station with largest mean flow. The below map function
  # uses the distance matrix from above to order 'clumped' (<= 20km) stations
  # by their most recent year of mean flow rate to produce a list of stations
  # to NOT label.
  stations_to_not_label = map(stations_with_labels$data, ~ {
    .x %>%
      #Add column identifying 'other' station for each distance measurement.
      mutate(other_station = stations_with_labels$STATION_NUMBER) %>% 
      #Add column identifying 'focal' station.
      mutate(this_station = .[.$dist_mat == 0,]$other_station) %>% 
      # Just keep stations within 20km
      filter(dist_mat <= 20000) %>% 
      # Add mean flow of last year for focal station
      mutate(focal_st_flow = stations_with_labels[stations_with_labels$STATION_NUMBER == .[.$dist_mat == 0,]$other_station,]$Average) %>% 
      # Add mean flow of last year for each nearby 'other station'
      left_join(stations_with_labels %>% st_drop_geometry() %>% dplyr::select(other_station = STATION_NUMBER, other_flow = Average),
                by = join_by(other_station)) %>% 
      # arrange (descending) by mean annual flow. Use this to find the largest 
      # station out of any 2+ stations within 20 km of each other.
      arrange(desc(other_flow)) %>% 
      # Drop the first row (the biggest station)
      slice(-1) %>% 
      dplyr::select(other_station) %>% 
      distinct()
  }) %>% 
    bind_rows() %>% 
    distinct(other_station)
  
  # Use the list of 'not-to-labels' to update our stations_marked table.
  stations_marked = stations_marked %>% 
    mutate(leaflet_label = ifelse(STATION_NUMBER %in% stations_to_not_label$other_station, NA, leaflet_label))
  
  # Determine the breaks/bins in circle size for plotting.
  stations_marked = stations_marked %>% 
    mutate(size_bins = cut(Average,number_size_classes,labels = F))
  
  size_legend_labels = tibble(
    range = cut(stations_marked$Average,number_size_classes, dig.lab = 4) %>% 
        levels() %>% 
        str_remove_all('\\(|\\]') %>% 
        str_replace_all(',',' - ') %>% 
        str_replace_all('^-[0-9\\.]*(?= )','0')
  ) %>% 
    mutate(bin_number = row_number()) %>% 
    deframe()
  
  # Create a colour palette to indicate the station of focus.
  mypal = colorFactor(palette = c("darkred", "darkblue"),
                   domain = stations_marked$focus_station,
                   ordered = F)
  
  # Find the station coordinates. Used in setting the map view.
  station_coords = st_coordinates(stations_marked %>% 
                                    filter(focus_station=='Station of Interest')) %>% as.data.frame

  
  main_leaf = leaflet(options = leafletOptions(zoomControl = FALSE)) %>% 
  addProviderTiles(leaflet_provider_choice) %>%
    removeControl(layerId = '') %>% 
    addSymbolsSize(
      lng = ~X,
      lat = ~Y,
      shape = 'circle',
      color = ~mypal(focus_station),
      fillColor = ~mypal(focus_station),
      fillOpacity = .5,
      baseSize = 30,
      values = ~size_bins,
      label = ~leaflet_label,
      labelOptions = labelOptions(noHide = T,
                                  direction = 'auto'),
      data = cbind(stations_marked, st_coordinates(stations_marked))
    ) %>%
    addLegendSize(
      title = HTML('Average Flow (m<sup>3</sup>/s)'),
      color = 'grey',
      values = (1:length(size_legend_labels)),
      breaks = size_legend_labels,
      baseSize = 45,
      shape = 'circle',
      data = stations_marked
    ) %>% 
    addMiniMap(
      tiles = providers$Wikimedia,
      position = 'topright',
      width = 200, 
      height = 200,
      toggleDisplay = FALSE
    ) %>% 
    setView(lng = station_coords$X, 
            lat = station_coords$Y, 
            zoom = 9) %>% 
    addScaleBar('bottomright') %>% 
    addLegend(
      title = 'Focal Station',
      pal = mypal,
      values = ~focus_station,
      position = 'topleft',
      data = stations_marked)
  
  # Find station region. Used in creating alt text.
  station_region = st_join(station,districts) %>% pull(DISTRICT_NAME)

  # Establish the name and filepath to leaflet plot.
  name_of_main_leaf = paste0('mainleaf_station_',this_station$STATION_NAME,'_',this_station$STATION_NUMBER,'.png')
  path_of_main_leaf = paste0(here::here(), "/print_ver/factsheet_figures/",name_of_main_leaf)
  
  #Save leaflet plot to disk. We load this back in with geom_bgimage.
  mapview::mapshot(x = main_leaf,
                   file = path_of_main_leaf,
                   vwidth = 1050, vheight = 650)
  
  ## 2. Leaflet map as a static shot.

  image_paths = data.frame(image = path_of_main_leaf)
  
  main_plot = ggplot(image_paths,aes(x=0.5,y=0.5)) +
    geom_bgimage(image = image_paths$image) +
    labs(title = paste0(
      this_station$STATION_NAME,
      " (",
      this_station$STATION_NUMBER,
      ")"
    )) +
    labs(x = "", y = "",
         alt = paste0("This station is located in the ",station_region,
                      " and there are ",
                      nrow(stations_marked[stations_marked$within_search_radius==T,]),
                      " stations within ",
                      station_search_dist/1000," kilometers.")) +
    theme(
      plot.title = element_markdown(face ='bold'),
      axis.ticks = element_blank(),
      axis.line = element_blank(),
      axis.text = element_blank()
    )
  
  return(main_plot)
}
```

```{r make_summary_table_function}
# Create Summary Table =============

make_summary_table = function(dat, monthly_dat, station){

  # Calculate the number of years included in the focal station's dataset. Also calculate
  # proportion of those years with 12 months of data.
  
  number_of_years_with_data = length(unique(monthly_dat$Year))
  total_possible_months = 12*number_of_years_with_data
  
  data_completeness = monthly_dat |> 
    count(Year) |> 
    summarise(data_completeness = 100 * (sum(n) / total_possible_months)) |>
    mutate(data_completeness = paste0(data_completeness,'%')) |> 
    pull(data_completeness)
  
  summary_most_recent_year = dat |>  
    mutate(`Years of Data` = paste0(min(Year),' to ',max(Year))) |> 
    mutate(`Data Completeness` = data_completeness) |> 
    slice_max(Year) |> 
    # Spatial join with natural resource districts.
    mutate(`Station Name` = station$STATION_NAME,
           `Station ID Number` = station$STATION_NUMBER,
           `Flow Group` = station$FlowGroup) %>% 
    mutate(Coordinates = paste0(round(st_coordinates(station),3),collapse = ', ')) %>% 
    mutate(Region = st_join(station,districts) %>% pull(DISTRICT_NAME)) %>% 
    # Count number of other stations within X km
    mutate(`Other Stations within x` = as.character( 
      nrow(
        st_join(
          st_buffer(
            station,dist= station_search_dist
            ), 
          stations,
          st_intersects)
        )-1 #The stations layer includes the focal station, so - 1.
      )
    ) %>% 
    mutate(across(c(Average,Min_7_Day), ~ paste0(round(.x,1),' m^3/s'))) %>% 
    dplyr::select(`Station Name`,`Station ID Number`,
                  `Years of Data`,`Data Completeness`,Coordinates,Region,
                  `Other Stations within x`,
                  Average,Min_7_Day) %>% 
    pivot_longer(cols = everything(), names_to = 'Parameter', values_to = 'Average Value') %>% 
    mutate(Parameter = case_when(
      Parameter == 'Median' ~ 'Median Flow',
      Parameter == 'Min_7_Day' ~ '7-Day Low Flow',
      Parameter == 'Other Stations within x' ~ paste0('Other Stations within ',station_search_dist/1000,' km'),
      T ~ Parameter
    )) %>% 
    setNames(c('Station Information',paste0('(',max(dat$Year),')'), ' '))
  
  summary_mry_raster = summary_most_recent_year %>% 
    flextable::flextable(cwidth = 2) %>%
    flextable::as_raster()
  
  summary_table_plot = ggplot() + 
    theme_void() + 
    annotation_custom(grid::rasterGrob(summary_mry_raster),
                      xmin = -Inf, xmax = Inf,
                      ymin = -Inf, ymax= Inf)
  
  return(summary_table_plot)
}
```

```{r make_linegraph_function}
## Line graph of flow variable over time

make_linegraph = function(dat, variable_to_plot) {

  label = variable_label_df[variable_label_df$varname == variable_to_plot,]$labelname 
  
  # Perform Mann-Kendall trend analysis on 2010 - present data.
  mk_results = mk_test(dat, variable_to_plot)
  
  # Join MK test results to the 2010-present day dataset.
  mk_line = dat %>%
    dplyr::select(Year,STATION_NUMBER,all_of(variable_to_plot)) %>% 
    cbind(mk_results) %>% 
    group_by(Year,STATION_NUMBER) %>% 
    summarise(SlopePreds = Intercept+Slope*Year)
  
  p = ggplot() + 
    geom_line(aes(y = SlopePreds,
                  x = Year),
              linewidth = 2,
              col = 'darkblue',
              alpha = 0.4,
              data = mk_line) +
    geom_point(aes(y = .data[[variable_to_plot]], 
                  x = Year),
              col = 'black',
              data = dat) +
    labs(y = paste0(label," (m<sup>3</sup>/s)"),
         caption = paste0(mk_results$trend_sig, " (P-value ~ ",round(mk_results$P_value,2),")")) +
    theme_minimal() +
    theme(axis.title.y = element_markdown(),
          legend.position = 'none')
  
  return(p)
}
```

```{r make_data_availability_histogram}
  ## Data Availability Histogram

make_data_histogram = function(dat){
  
  data_histogram = dat %>% 
    count(Year, name = 'number_months') %>% 
    mutate(number_months = 12 - number_months) %>% 
    ggplot() + 
    geom_col(aes(x = Year, y = number_months),
           fill = '#BE3131', col = 'transparent') + 
    geom_label_repel(aes(x = Year, y = number_months,
                         label = Year),
                     data = . %>% filter(number_months > 0),
                     nudge_y = 1) + 
    scale_y_continuous(breaks = c(0,4,8,12), limits = c(0,12)) +
    labs(y = "Months Without Data") +
    theme_minimal()

  return(data_histogram)
}
```

```{r make_hydrograph}

make_hydrograph = function(this_station){
  
  station_name = this_station$STATION_NAME
  
  # Pull data from tidyhydat database for our station of interest; calculate quantiles.
  daily_flows = hy_daily_flows(this_station$STATION_NUMBER) |> 
    filter(!is.na(Value)) |> 
    mutate(Day = lubridate::day(Date),
           Month = lubridate::month(Date)) |> 
    group_by(Month, Day, STATION_NUMBER) %>%
    reframe(percentiles = list(quantile(Value, probs = c(0.05,0.25,0.50,0.75,0.95)))) %>%
    unnest_wider(percentiles) |>
    dplyr::rename(five_perc = `5%`, twentyfive_perc = `25%`,
                  median_flow = `50%`,
                  seventyfive_perc = `75%`, ninetyfive_perc = `95%`)
  
  # Convert the month variable into an ordered factor such that
  # October is the first year on the X axis (i.e. we're using the 'water calendar')
  plotting_df = daily_flows %>%
    ## Convert from calendar year to 'water year'
    # Step 1. Add in 'dummy year' to facilitate plotting x axis...
    mutate(dummy_year = 2000) |> 
    mutate(plotting_date = lubridate::ymd(paste(dummy_year,Month,Day,sep = '-'))) |> 
    mutate(Month = month.abb[Month]) |>
    mutate(month_label = factor(Month, levels = c(month.abb[10:12],month.abb[1:9]))) %>%
    arrange(month_label) |>
    # Add labels for the ribbons we'll add to the figure.
    mutate(median_line_label = 'Median Flow') %>%
    mutate(fifty_pct_label = '"Normal" range (50%) of flow') %>%
    mutate(ninety_pct_label = 'Range of 90% of flow')
  
  p = plotting_df %>%
    ggplot() +
    geom_ribbon(aes(x = plotting_date, ymin = five_perc, ymax = ninetyfive_perc, fill = ninety_pct_label)) +
    geom_ribbon(aes(x = plotting_date, ymin = twentyfive_perc, ymax = seventyfive_perc, fill = fifty_pct_label)) +
    geom_line(aes(x = plotting_date, y = median_flow, colour = median_line_label),
              linewidth = 1) +
    scale_colour_manual(values = c("Median Flow" = "#2d7ca1")) +
    scale_fill_manual(values = c("Range of 90% of flow" = "#ceeaed",
                                 '"Normal" range (50%) of flow' = 'lightblue')) +
    scale_x_date(date_breaks = '1 month', date_labels = '%b') +
    # scale_x_continuous(breaks = c(1:12),
    #                    labels = plotting_df$month_label[c(1:12)]) +
    labs(y = 'Average Discharge (m<sup>3</sup>/s)',
         x = '',
         title = '*Daily Stream or River Discharge*',
         subtitle = station_name,
         col = '',
         fill = '') +
    theme(axis.title.y = element_markdown(size = 15),
          axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12),
          legend.position = 'top',
          plot.title = element_markdown(hjust = 0.5),
          panel.background = element_rect(fill = 'transparent'),
          panel.grid.major = element_line(colour = 'grey'))
  
  return(p)
}
```
  
```{r make_circle_plot_function}
# make_circle_plot = function(dat,station,variable_to_plot){
# 
#   label = variable_label_df[variable_label_df$varname == variable_to_plot,]$labelname
# 
#   plot_dat = dat %>%
#     filter(Month != 'All') %>%
#     rowwise() %>%
#     mutate(Month = which(month.abb == Month)) %>%
#     as_tibble() %>%
#     mutate(Date = lubridate::ymd(paste0(Year,'-',Month,'-01'))) %>%
#     rename("Value" = variable_to_plot)
# 
#   log10_floor <- function(x) {
#     10^(floor(log10(x)))
#   }
# 
#   log10_ceiling <- function(x) {
#     10^(ceiling(log10(x)))
#   }
# 
#   flow_var_minimum = min(plot_dat$Value)
#   flow_var_min_mag = log10_floor(flow_var_minimum)
#   flow_var_maximum = max(plot_dat$Value)
#   flow_var_max_mag = log10_ceiling(flow_var_maximum)
# 
#   y_axis_text_labels = data.frame(
#     y_values = log10(flow_var_min_mag):log10(flow_var_max_mag)
#   ) %>%
#     mutate(
#       labels = 10^y_values) %>%
#     mutate(Month = 11)
# 
#   x_axis_text_labels = data.frame(
#     x_values = c(3,6,9,12),
#     y_values = rep(log10(flow_var_max_mag),4),
#     labels = c("Spring",
#                "Summer",
#                "Autumn",
#                "Winter")
#   )
# 
#   plot_dat %>%
#     mutate(Date = as.factor(Date)) %>%
#     mutate(Date = forcats::fct_inorder(Date)) %>%
#     ggplot(aes(x = Month, y = log10(Value))) +
#     geom_line(aes(colour = Year, group = Year),
#               linewidth = 1.5,
#               alpha = 0.8
#     ) +
#     geom_label(aes(x = Month, y = y_values, label = labels),
#                alpha = 0.6,
#                data = y_axis_text_labels) +
#     geom_label(aes(x = x_values, y = y_values, label = labels),
#                alpha = 0.6,
#                data = x_axis_text_labels) +
#     labs(caption = paste0(label," (m<sup>3</sup>/s) by Month of Year")) +
#     coord_polar() +
#     scale_colour_distiller(palette = 'Spectral') +
#     scale_y_continuous(limits = c(1,5)) +
#     theme(
#       plot.background = element_rect(fill = '#302e27'),
#       panel.background = element_rect(fill = '#302e27'),
#       axis.title = element_blank(),
#       axis.text = element_blank(),
#       plot.caption = element_markdown(colour = 'white',
#                                       size = 12,
#                                       face = 'bold',
#                                       # halign = 0.5,
#                                       hjust = 0.5)
#       )
# }
```

```{r make_day_of_year_circle_plot}

make_day_of_year_circle_plot = function(dat,variable_to_plot){

  label = variable_label_df[variable_label_df$varname == variable_to_plot,]$labelname 
  
  plot_dat = this_dat %>%
    filter(!is.na(variable_to_plot))
  
  max_year = max(plot_dat$Year)
  min_year = min(plot_dat$Year)
  
  y_axis_text_labels = tibble(
    # Use 'floor division' (name?) to create prettier, round-number breaks (i.e. 1940 instead of 1938)
    y = 10*seq(min_year+20, max_year, by = (max_year-min_year)/4) %/% 10
  ) %>% 
    mutate(x = 300)
  
  x_axis_text_labels = data.frame(
    x_values = c(80,172,265,355),
    y_values = rep(2030,4),
    labels = c("Spring",
               "Summer",
               "Autumn",
               "Winter")
  )
  
  plot_dat %>% 
    ggplot(aes(x = .data[[variable_to_plot]], y = Year)) +
    geom_point(aes(colour = Year),alpha = 0.8) +
    geom_text(aes(x = x_values, y = y_values, label = labels),
               alpha = 0.6,
               data = x_axis_text_labels) +
    geom_text(aes(x = x, y = y, label = y),
              col = '#706c5e',
               data = y_axis_text_labels) + 
    labs(title = label) +
    coord_polar() +
    scale_colour_distiller(palette = 'Spectral') +
    labs(y = "") +
    scale_x_continuous(limits = c(0,365)) +
    scale_y_continuous(limits = c(min_year,max_year+10)) +
    theme(
      # plot.title = element_text(colour = 'white', hjust = 0.5),
      # plot.background = element_rect(fill = '#302e27'),
      # panel.background = element_rect(fill = '#302e27'),
      plot.title = element_text(hjust = 0.5),
      panel.grid = element_line(colour = 'darkgrey'),
      plot.background = element_rect(fill = 'white'),
      panel.background = element_rect(fill = 'white'),
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank()
    )
}
```

# **Long-term Change in Timing & Volume of River Flow in B.C.**

Changes in the timing and volume of river flow can affect both natural ecosystems and human communities---they can affect our ability to predict and manage seasonal water resources and flood risks, and may impact natural systems that rely on natural flow patterns. Low river flows in summer may reduce the amount of water available for agriculture, hydroelectric power generation, industry and communities. Low river flows are also associated with declines in water quality and warmer water temperatures which can threaten the health of aquatic ecosystems. This indicator measures changes in timing and volume of river flow across the province over a `r max(annual_flow_dat$Year)-min(annual_flow_dat$Year)`-year period (`r oldest_data` to `r most_recent_data`).

-   **Long-term change in magnitude or timing of river flow has occurred at several locations in B.C.** The magnitude of flow is defined as the mean annual flow in cubic meters per second and the timing of flow is defined as the day of year by which one-half of the total annual volume of each river has passed.

-   **Many factors affect trends in timing and volume of river flow.** Changes in [temperature](https://www.env.gov.bc.ca/soe/indicators/climate-change/temp.html), [precipitation](https://www.env.gov.bc.ca/soe/indicators/climate-change/precip.html), and evapotranspiration can all affect the timing and volume of river flow. Trends also depend on the location, size, elevation, and whether the river basin is predominately rainfall- or snowmelt-fed---and whether substantial glacial cover exists. Thus, trends in river timing and flows are not uniform across B.C.

```{r compose_factsheet_pages, fig.height=12, fig.width=8}

stations_all = stations

stations = stations_all[c(1:10),]

my_progress_bar = knitrProgressBar::progress_estimated(n = nrow(stations))

# for(i in 1:nrow(stations)){
for(i in 1:1){

  # Update progress bar (helps us know how far we've gotten!)
  my_progress_bar$tick()
  my_progress_bar$print()
  
  # Subset all data for this station ===========================================
  
  this_station = stations %>% 
    filter(row_number() == i)
  
  this_dat_annual = annual_flow_dat |> 
    filter(STATION_NUMBER == this_station$STATION_NUMBER)

  this_dat_monthly = monthly_flow_dat |> 
    filter(STATION_NUMBER == this_station$STATION_NUMBER)
  
  this_hydrograph_dat = hy_daily_flows(this_station$STATION_NUMBER)
  
  # Create page elements =======================================================
  
  main_plot = make_leaflet(station = this_station, stations = stations)

  summary_table_plot = make_summary_table(dat = this_dat_annual, monthly_dat = this_dat_monthly, station = this_station)

  average_flow_linegraph = make_linegraph(dat = this_dat_annual, variable_to_plot = 'Average')
  
  lowflow_sevenday_linegraph = make_linegraph(dat = this_dat_annual, variable_to_plot = 'Min_7_Day')
  
  # dat_year_histogram = make_data_histogram(this_dat_monthly)
  
  hydrograph_plot = make_hydrograph(this_station)
  
  min_flow_plot = make_linegraph(this_dat_annual, variable_to_plot = 'Min_7_Day_DoY')
  
  half_flow_plot = make_linegraph(this_dat_annual, variable_to_plot = 'DoY_50pct_TotalQ')
   
  # Compose elements===================================
  
  col_one = ggarrange(
    summary_table_plot
  )
  
  col_two = ggarrange(
    average_flow_linegraph,
    lowflow_sevenday_linegraph,
    nrow = 2
  )
  
  # Put it all together ========================================================

  this_arranged_plot = ggarrange(
    main_plot,
    ggarrange(
      col_one,
      col_two,
      ncol = 2
    ),
    hydrograph_plot,
    nrow = 3,
    heights = c(5,4,3)
  )
  
  plot(this_arranged_plot)
}
```

```{r}

```

\newpage

## Data

\*By accessing these datasets, you agree to the licence associated with each file, as indicated below.

------------------------------------------------------------------------

\newpage
