---
topic: "water"
title: "Long-Term Trends in River Flows in B.C."
output: envreportutils.internal::print_ver
---
<!--
Copyright 2024 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->

```{r setup, echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}

library(dplyr)
library(ggplot2)
library(grid)
library(RColorBrewer)
library(scales)
library(bcgroundwater)
library(envreportutils)
library(gridExtra)
library(xtable)
library(lubridate)
library(forcats)
library(here)
library(magick)
library(png)
library(cowplot)
library(here)
library(sf)
library(stringr)

knitr::opts_chunk$set(cache = FALSE)

theme_set(theme_classic() +
            theme(text = element_text(colour = "black"),
                  axis.line = element_blank(),
                  axis.ticks = element_blank(),
                  panel.grid.major = element_line(colour = "grey85", size = 0.5,
                                                linetype = 1),
                  panel.grid.minor = element_line(colour = "grey90", size = 0.5,
                                                linetype = 1),
                  panel.grid.major.x = element_blank(),
                  panel.spacing = unit(0.6, "lines"),
                  plot.title = element_text(vjust = 2, hjust = 0.5),
                  axis.title = element_text(vjust = 0.1),
                  legend.position = "bottom", legend.title = element_blank(),
                  legend.text = element_text(size = 12),
                  axis.text.x = element_blank(),
                  strip.background = element_blank()))

label.colour <- "black" 
colour.scale <- brewer.pal(3,"Blues")

updateDate <- "February 2024"

annual_flow_dat = readRDS(here("app/www","annual_flow_dat.rds"))
stations_sf = st_read(here("app/www","stations.gpkg"))
mk_results_tbl = readRDS("print_ver/out/mk_results_tbl.rds")

recent_stations = annual_flow_dat %>%
  group_by(STATION_NUMBER) %>%
  summarise(minYear = min(Year)) %>%
  filter(minYear>=1992) %>%
  pull(STATION_NUMBER)

stations_sf_filtered = stations_sf %>% 
  filter(! (is.na(keep))) #%>%
  # filter(! STATION_NUMBER %in% recent_stations)

# summary table of number of stations in each category
summary_table_cat = mk_results_tbl %>%
  filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Var, category) %>%
  summarise(n = n())


summary_table_direction = summary_table_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Var, direction) %>%
  summarise(n = sum(n))

#Subset Average
average_cat = summary_table_cat %>%
  filter(Var == "Average")

average_direction = summary_table_direction %>%
  filter(Var == "Average")

#Subset Peak Flow
peak_cat = summary_table_cat %>%
  filter(Var == "Peak Flow")

peak_direction = summary_table_direction %>%
  filter(Var == "Peak Flow")

#Subset Low Flow
Low_cat = summary_table_cat %>%
  filter(Var == "Low Summer Flow")

Low_direction = summary_table_direction %>%
  filter(Var == "Low Summer Flow")

#Subset Freshet
freshet_cat = summary_table_cat %>%
  filter(Var == "Date of Freshet")

freshet_direction = summary_table_direction %>%
  filter(Var == "Date of Freshet")

#Subset Freshet
low_date_cat = summary_table_cat %>%
  filter(Var == "Date of Low Summer Flow")

low_date_direction = summary_table_direction %>%
  filter(Var == "Date of Low Summer Flow")

load(here("print_ver/out","figures.Rdata"))
```

Changes in the timing and volume of river flow can affect both natural ecosystems and human 
communities - they can affect our ability to predict and manage seasonal water resources and
flood risks, and may impact natural systems that rely on natural flow patterns. Low river flows
in summer may reduce the amount of water available for agriculture, hydroelectric power generation,
industry and communities. Low river flows are also associated with declines in water quality and 
warmer water temperatures which can threaten the health of aquatic ecosystems. The indicator measures
several metrics representing river flow at `r nrow(stations_sf_filtered)` stations (`r nrow(stations_sf)`
when including upstream stations) across the Province. These stations have been monitored for a minimum of 
`r min(stations_sf_filtered$N_years)` years and a maximum of `r max(stations_sf_filtered$N_years)` years 
between `r min(stations_sf_filtered$Min_Year)` and `r max(stations_sf_filtered$Max_Year)`.

# Description of metrics

## Average annual flow

Average annual flow was calculated within each water year (October - October). The mean daily flow was identified for each station and each year of available data, and
the trend was estimated. We removed years that had more than 30% daily missing data to avoid potentially biased annual measures of average flow.

## Volume of peak flow and timing of freshet

Volume of peak flow and timing of freshet were calculated using water year (October - October). Volume of peak flow was calculated using a 3-day rolling window, identifying the maximum water volume measured at each station for each year. Freshet is defined as the flood of a river, often due to heavy rain or snow melt. For this analysis, we define date of freshet as the date whereby 50% of river flow for that water year has taken place. 

## Date and volume of low flow

Timing and volume of low flows was calculated based on the low flow year (April - April). Volume of low flow was estimated based on a 7-day rolling window across the summer (May - October), and the minimum volume was identified for each year at each station. Timing of low flow was calculated based on the Mean Annual Discharge (MAD; ref). The first date when 50% of MAD was reached after freshet was used as the definition of the beginning of summer low flows. 

\vspace{.2cm}

# Trend Results

```{r fig.width = 10, fig.height = 8, echo = FALSE}
mk_all
```

## Flow volume

The metrics measured at each station can be broken down into two types: change in the magnitude of river flow and change in the timing of river flow. Of the `r nrow(stations_sf_filtered)` stations monitored across the Province, `r average_direction$n[average_direction$direction == "positive"]` stations identified increasing flows, of which `r average_cat$n[average_cat$category == "> 0.5% increase"]` were increasing by more than 0.5% per year. In contrast,  `r average_direction$n[average_direction$direction == "negative"]` displayed decreasing flows, while the remaining `r average_direction$n[average_direction$direction == "no change"]` were either less than 0.1% a year or had non-significant trends.

Similarly, peak flows trends were almost evenly split between increasing (`r peak_direction$n[peak_direction$direction == "positive"]` stations) and decreasing (`r peak_direction$n[peak_direction$direction == "negative"]` stations), while the vast majority were found to have little significant change (`r peak_direction$n[peak_direction$direction == "no change"]` stations).

Low summer flow volume has been significantly decreasing over time for a substantial number of rivers. `r Low_direction$n[Low_direction$direction == "negative"]` stations had decreasing summer flows, and `r Low_cat$n[Low_cat$category == "> 0.5% decrease"]` were greater than 0.5% per year.

## Flow timing

Date of freshet has been occurring gradually earlier for a large number of locations. `r freshet_direction$n[freshet_direction$direction == "negative"]` have been trending earlier, and `r freshet_cat$n[freshet_cat$category == "> 0.2 days earlier"]` are occurring 0.2 days earlier per year. Only `r freshet_direction$n[freshet_direction$direction == "positive"]` was trending later, while `r freshet_direction$n[freshet_direction$direction == "no change"] + freshet_cat$n[freshet_cat$category == "< 0.1 days change"]` displayed trends that were less than 0.1 days change or not significant.

The majority of locations did not show significant trends for timing of low flow. Of those that did, more stations had low flows occurring earlier (`r low_date_direction$n[low_date_direction$direction == "negative"]`) than later (`r low_date_direction$n[low_date_direction$direction == "positive"]`).

# Changes in Flow Volume by Month

## Median Monthly flows

```{r fig.width = 10, fig.height=8, echo = FALSE}
monthly_average_bar_plot
```

Changes in average river flow across years varies by month. 

## Monthly low flows
```{r fig.width = 10, fig.height=8, echo = FALSE}
monthly_low_flow_bar_plot
```

# Changes in Volume and Timing by Hydrologic Zones

# Flow volume

## Average annual flow
```{r fig.width = 10, fig.height=8, echo = FALSE}
annual_bar_plot
```

## Peak flow
```{r fig.width = 10, fig.height=8, echo = FALSE}
peak_bar_plot
```

## Low flow
```{r fig.width = 10, fig.height=8, echo = FALSE}
low_bar_plot
```

# Flow timing

## Date of freshet
```{r fig.width = 10, fig.height=8, echo = FALSE}
freshet_bar_plot
```

## Date of low flow
```{r fig.width = 10, fig.height=8, echo = FALSE}
date_low_bar_plot
```

## Useful Links

## Data

The data sets are all sourced from the B.C. Data Catalogue, distributed under the [Open Government Licence - British Columbia](https://www2.gov.bc.ca/gov/content?id=A519A56BC2BF44E4A008B33FCF527F61).

-   [Indicator Summary Data: Long-term Trends in Timing and Volume of B/C/s River Flow](https://catalogue.data.gov.bc.ca/dataset/indicator-summary-data-change-in-timing-and-volume-of-river-flow-in-bc-1912-2012-)

- River Flow Stations and daily flow data accessed with ['tidyhydat' R Package](https://cran.r-project.org/web/packages/tidyhydat/index.html)

- [Hydrologic Zones](https://catalogue.data.gov.bc.ca/dataset/bc-hydrologic-zones).

----

Published and Available Online at Environmental Reporting BC (`r updateDate`): <https://www.env.gov.bc.ca/soe/indicators/climate-change/rivers.html0>

Email correspondence to: envreportbc@gov.bc.ca

*Suggested Citation*:  
Environmental Reporting BC. 2024.  Long-Term Trends in River Flows in B.C. State of Environment Reporting, Ministry of Environment and Climate Change Strategy, British Columbia, Canada.

## Methods

### Filtering procedure

To account for variation in monitoring at different stations on different rivers, a set of filtering procedures were implemented to reduce potential for erroneous trend estimations. These included:

- **Data gaps**: Monitoring stations containing gaps of 5 or more years had all annual data preceding the gap removed, to maintain relatively continuous datasets at each station for trend analyses.

- **Recently active**: Stations were required to be active as recently as 2019 for inclusion in the analysis.

- **Minimum sample size**: Only stations with more than 10 years of monitoring data were retained.

- **Regulated Rivers**: A river is defined as regulated if the flow regime has been modified and results in a 10% increase or decrease in flow (REF). For the purpose of this 
analysis, regulated stations were removed from the analysis (apart from certain instances where flow was deemed close to natural).

- **Upstream stations**: A number of rivers include multiple monitoring stations. In such situations, only the downstream stations are included (although upstream locations can be inspected using the online app).

- **Recently installed stations**: Stations that have been installed within the last 30 years were...... FINISH

### References


